---
title: "Assignment 3"
author: "Kai Zhang, Sherida van den Bent, Chang Liu"
date: "2020/03/22"
output: pdf_document
---
```{r echo=FALSE,,message=FALSE}
set.seed(1234)
```
## Exercise 1
**a)**
```{r}
fruitflies = read.table('fruitflies.txt', header=TRUE)
fruitflies$loglongevity = log(fruitflies$longevity)
plot(loglongevity~thorax,pch=unclass(activity), data=fruitflies)
for (i in c('high', 'low', 'isolated')) abline(lm(loglongevity~thorax,data=fruitflies[fruitflies$activity==i,]))

par(mfrow=c(1,2))
boxplot(loglongevity~activity, data=fruitflies); boxplot(loglongevity~thorax, data=fruitflies)
fruitfliesaov = lm(loglongevity~activity, data=fruitflies)
anova(fruitfliesaov); summary(fruitfliesaov)
plot(fruitfliesaov, 2); plot(fruitfliesaov, 1)
shapiro.test(residuals(object = fruitfliesaov))
```
From the first plot we can see that loglongevity clearly increases with thorax. Its dependence on activity is not so clear. From the two box plots, it is clear that both activity and thorax have effect on loglongevity.
Then we use ANOVA to test $H_{0}:\mu_{high}=\mu_{isolated}=\mu_{low}$. p-value = 1.798e-07 < 0.05. Conclusion we reject null hypothesis and sexual activity do influences longevity.
From the summary, we get $\hat{\mu}=3.60212$, $\hat{\alpha}_{isolated}=0.51722$, $\hat{\alpha}_{low}=0.39771$. So estimated loglongevities for high is 3.60212, for isolated is 3.60212+0.51722=4.11934 and for low is 3.60212+0.39771=3.99983. And estimated longevities for high is 36.67591, for isolated is 61.51863 and for low is 3.60212+0.39771=54.58887.
We check the assumption of ANOVA and found Normal QQ-plot looks like has a curve, and Shapiro-Wilk normality test gives p-vlaue=0.008652, it means the conclusion is not reliable.

**b)**
```{r}
fruitfliesaov = lm(loglongevity~thorax+activity, data=fruitflies)
anova(fruitfliesaov); summary(fruitfliesaov)
par(mfrow=c(1,2))
plot(fruitfliesaov, 2); plot(fruitfliesaov, 1)
mean(fruitflies$thorax)
shapiro.test(residuals(object = fruitfliesaov))
```
We use ANCOVA to test $H_{0}:\mu_{high}=\mu_{isolated}=\mu_{low}$. p-value = 4e-09 < 0.05. Conclusion we reject null hypothesis and sexual activity do influences longevity.
From the summary, we get $\hat{\mu}=1.21893$, $\hat{\alpha}_{isolated}=0.40998$, $\hat{\alpha}_{low}=0.28570$. So sexual activity decrease longevity
We check the assumption of ANOVA and found Normal QQ-plot looks like has a curve, it means the conclusion is not reliable.
The model is $\hat{Y_{i}}=\hat{\mu}+\hat{\alpha_{i}}+\hat{\beta} X_{i}$.
The average thorax length is 0.8245333. 
The estimated loglongevity for a fly with average thorax length in isolated: $\hat{Y_{isolated}}=1.21893+0.40998+2.97899*0.8245333=4.08518645537$. 
For low:$\hat{Y_{low}}=1.21893+0.28570+2.97899*0.8245333=3.96090645537$.
For High:$\hat{Y_{high}}=1.21893+0+2.97899*0.8245333=3.67520645537$.
So longevity for isolated is $e^{Y_{isolated}} = e^{4.09}=59.45$.
Longevity for low is $e^{Y_{low}} = e^{3.96}=52.50$.
Longevity for high is $e^{Y_{high}} = e^{3.68}=39.46$.
We check the assumption of ANCOVA. After creating a QQ-plot, we cannot tell whether it is a normal distribution because of some outliers. Therefore we also use the Shapiro-Wilk normality test. Based on the p-value(0.05748 > 0.05), we could say that it is probably normally distributed. So the first assumption of normal distribution of the model residuals has been met. The fitted plot seems to indicate that the residuals and the fitted values are uncorrelated. We can rely on the ANCOVA test.

**c)**
```{r}
fruitfliesaov = lm(loglongevity~activity*thorax, data=fruitflies)
summary(fruitfliesaov)
```
From the plot in a), it is clear that the three lines would be parallel. P-value for $H_0:\mu_{isolated}=\mu_{high}$ and $H_0:\mu_{low}=\mu_{high}$ are 0.0545 and 0.2771. So we do not reject $H_0$. Conclusion: this dependence is similar under all three conditions of sexual activity.

**d)**
I prefer the one with thorax length. Because the one without thorax length doesn't fit assumption of ANOVA. 

**e)**
From the plot from b), we can say the assumption of normal distribution of the model residuals and homogeneity of variance of the groups has been met.

**f)**
```{r}
fruitfliesaov = lm(longevity~thorax+activity, data=fruitflies)
anova(fruitfliesaov); summary(fruitfliesaov)
par(mfrow=c(1,2))
plot(fruitfliesaov, 2); plot(fruitfliesaov, 1)
```
From the fitted plot, we can see a certain pattern in residuals fitted plot. So the model with longevity instead of loglongevity is not reliable.

## Exercise 2
```{r}
require(ggplot2)
require(plyr)
require(dplyr)
require(aod)
par(mfrow = c(1, 1))
```
**a)**
```{r}
grades = read.table('psi.txt')
head(grades)
```
The table psi consists of three columns; the boolean 'passed' signifying if the student passed the assignment or not, the boolean 'psi' signifying if the student was taught by the regular method (0) or if the student was taught by the new psi method (1). The third column is gpa, signifying the average grade of the student on a 4-point scale.

```{r}
summary(grades)
```
Above you will find the complete summary in a table.

The barplots below can already give us an inkling how well the psi method performed relatively to the regular teaching method. This barplot shows the frequency of passes and fails from the students getting regular teaching methods vs psi teaching methods.

```{r}
all_psi = subset(grades, psi == 1)
all_reg = subset(grades, psi == 0)
par(mfrow = c(1, 2))
barplot(table(all_reg$passed),main='Regular',names.arg=c('fail','pass'))
barplot(table(all_psi$passed),main='Psi',names.arg=c('fail','pass'))
```


**b)**
```{r}
sapply(grades, sd)
xtabs(~passed + psi, data = grades)

grades$psi <- factor(grades$psi)
mylogit <- glm(passed ~ gpa + psi, data = grades, family = "binomial")
summary(mylogit)
```

Psi works; the coefficient of psi1 is statistically significant, and changes the log chances of passing by 2.338. This is quite a lot; PSI is a lot better than the regular teaching methods.

**c)**
```{r}
testdata <- with(grades, data.frame(gpa = 3, psi = factor(0:1)))
testdata$passed <- predict(mylogit, newdata = testdata, type = "response")
testdata
```

The predicted probability of a student with a gpa of 3, and getting the psi teaching method is 0.48158645, and the same student getting the regular teaching method has a passing probability of 0.08230274.

**d)**

As said in question b, psi changes the log chances of passing by 2.338. This is irrellevant of gpa score.

**e)**
```{r}
x=matrix(c(3,15,8,6),2,2)
x
```

The numbers 15 and 6 represent the students that did not show any improvement. 15 is the number of kids without psi that didn't show any improvement, and 6 is the number of kids with psi that didn't show any improvement.

```{r}
fisher.test(x)
```
The p-value is 0.0265; this is lower than 0.05, so H0 is rejected. This means that p1 != p1; the odds of improvement by using psi and by using the regular teaching methods are not equal.

# F
Yes, the second aproach is wrong. It doesn't take the gpa of the students into account (it isn't even in the data) so the outcome could be greatly skewed by the gpa of the students.

# G

Logistic regression:
advantage: Can predict new outcomes based on new data
disadvantage: Needs lots of data to function properly
  
Fisher test:
advantage: Very easy to add extra testcases without getting too big of a table
disadvantage: Doesn't take gpa into account

#3 Read in data
```{r}
africa=read.table("africa.txt",header=TRUE);attach(africa)
africa
```
## 3.a)
```{r}
par(mfrow=c(3,3));
for (i in c(2,3,4,5,6,7,8,9)) hist(africa[,i],main=names(africa)[i])
plot(africa);
```
From the hist chart and the plot chart we can see that it also has collinearity problem amongst explanatory variables.
```{r}
africa_glm = glm(miltcoup~oligarchy+pollib+parties+pctvote+popn+size+numelec+numregim,family = poisson)
summary(africa_glm)
```
```{r}
confint(africa_glm)
```
```{r}
coef(africa_glm)
```
The Poisson-regression's result is: $\log(\lambda)=0.46070 + 0.1325623720oligarchy + -1.2275514078pollib+0.0627986634parties$
## 3b)
```{r}
africa_glm = glm(miltcoup~oligarchy+pollib+parties+pctvote+popn+size+numelec+numregim,family = poisson)
summary(africa_glm)
```
|           |Estimate Std. |Error      |z value |Pr(>|z|)  |        |
|-----------|--------------|-----------|--------|----------|--------|   
|(Intercept)|-0.5102693    |0.9053301  |-0.564  |0.57301   |        |
|oligarchy  | 0.0730814    |0.0345958  | 2.112  |0.03465 * |        |
|pollib     |-0.7129779    |0.2725635  |-2.616  |0.00890 **|        |
|parties    | 0.0307739    |0.0111873  | 2.751  |0.00595 **|        |
|pctvote    | 0.0138722    |0.0097526  | 1.422  |0.15491   |        |
|popn       | 0.0093429    |0.0065950  | 1.417  |0.15658   |        |
|size       |-0.0001900    |0.0002485  |-0.765  |0.44447   |        |
|numelec    |-0.0160783    |0.0654842  |-0.246  |0.80605   |=>remove|
|numregim   | 0.1917349    |0.2292890  | 0.836  |0.40303   |        |
```{r}
africa_glm = glm(miltcoup~oligarchy+pollib+parties+pctvote+popn+size+numregim,family = poisson)
summary(africa_glm)
```
|           |Estimate Std. |Error      |z value |Pr(>|z|)  |        |
|-----------|--------------|-----------|--------|----------|--------|   
|(Intercept)|-0.6078028    |0.8239267  |-0.738  |0.46070   |        |
|oligarchy  | 0.0781368    |0.0277656  | 2.814  |0.00489 **|        |
|pollib     |-0.6773897    |0.2290130  |-2.958  |0.00310 **|        |
|parties    | 0.0296786    |0.0102888  | 2.885  |0.00392 **|        |
|pctvote    | 0.0131290    |0.0092895  | 1.413  |0.15756   |        |
|popn       | 0.0089313    |0.0063746  | 1.401  |0.16120   |        |
|size       |-0.0002021    |0.0002436  |-0.830  |0.40682   |        |
|numregim   | 0.1758198    |0.2210498  | 0.795  |0.42639   |=>remove|
```{r}
africa_glm = glm(miltcoup~oligarchy+pollib+parties+pctvote+popn+size,family = poisson);summary(africa_glm)
```
|           |Estimate Std. |Error      |z value |Pr(>|z|)    |        |
|-----------|--------------|-----------|--------|------------|--------|
|(Intercept)|-0.1126871    |0.5163030  |-0.218  |0.827228    |        |
|oligarchy  | 0.0859620    |0.0259100  | 3.318  |0.000908 ***|        |
|pollib     |-0.6894029    |0.2278572  |-3.026  |0.002481 ** |        |
|parties    | 0.0291944    |0.0101954  | 2.863  |0.004190 ** |        |
|pctvote    | 0.0141588    |0.0091980  | 1.539  |0.123723    |        |
|popn       | 0.0062736    |0.0053994  | 1.162  |0.245272    |        |
|size       |-0.0001950    |0.0002425  |-0.804  |0.421378    |=>remove|
```{r}
africa_glm = glm(miltcoup~oligarchy+pollib+parties+pctvote+popn,family = poisson);summary(africa_glm)
```
|           |Estimate Std. |Error      |z value |Pr(>|z|)    |        |
|-----------|--------------|-----------|--------|------------|--------|
|(Intercept)|-0.244466     |0.495708   |-0.493  |0.62190     |        |
|oligarchy  | 0.083168     |0.025437   | 3.270  |0.00108 **  |        |
|pollib     |-0.652830     |0.221234   |-2.951  |0.00317 **  |        |
|parties    | 0.029800     |0.010294   | 2.895  |0.00379 **  |        |
|pctvote    | 0.013842     |0.009282   | 1.491  |0.13591     |        |
|popn       | 0.005587     |0.005378   | 1.039  |0.29883     |=>remove|
```{r}
africa_glm = glm(miltcoup~oligarchy+pollib+parties+pctvote,family = poisson);summary(africa_glm)
```
|           |Estimate Std. |Error      |z value |Pr(>|z|)    |        |
|-----------|--------------|-----------|--------|------------|--------| 
|(Intercept)|-0.093657     |0.463279   |-0.202  | 0.83979    |        |
|oligarchy  | 0.095358     |0.022421   | 4.253  |2.11e-05 ***|        |
|pollib     |-0.666615     |0.217564   |-3.064  | 0.00218 ** |        |
|parties    | 0.025630     |0.009502   | 2.697  | 0.00699 ** |        |
|pctvote    | 0.012134     |0.009056   | 1.340  | 0.18031    |=>remove| 
```{r}
africa_glm = glm(miltcoup~oligarchy+pollib+parties,family = poisson);summary(africa_glm)
```
|           |Estimate Std. |Error      |z value |Pr(>|z|)    |        |
|-----------|--------------|-----------|--------|------------|--------|   
|(Intercept)|  0.251377    |0.372689   | 0.674  | 0.50000    |        |
|oligarchy  |  0.092622    |0.021779   | 4.253  |2.11e-05 ***|        |
|pollib     | -0.574103    |0.204383   |-2.809  | 0.00497 ** |        |
|parties    |  0.022059    |0.008955   | 2.463  | 0.01377 *  |        |
Finished! Nothing to remove!
```{r}
confint(africa_glm);
```
```{r}
plot(log(fitted(africa_glm)),residuals(africa_glm))
```
This plot seems ok, there's not specific structure inside.
```{r}
plot(log(fitted(africa_glm)),residuals(africa_glm,type="response"))
```
The response residuals clearly increase with the (logarithm) of the fitted values, as expected under a Poisson model.
```{r}
coef(africa_glm)
```
So the final result is: $\log(\lambda) = 0.25137715+0.09262231oligarchy-0.57410330pollib+0.02205864parties$
